Fractional Types

### 2016-12-19 typespp.tex
typo
missing '['
sec 5 changes

Each combinator $\rho_\bullet$ has an inverse. A rule $\mathbb{T_1}_\bullet \leftrightblackspoon

@@

Compact closed categories have various interesting notions, such as
duals, name and coname, which eventually lead to a certain notion of
higher order functions.  Here we show that the first three (duals,
name and coname) are definable circuits in $\Pi^/$.

\subsection{\texttt{dual}}

@@

\subsection{\texttt{name} and \texttt{coname}}

We also show circuits that construct the \texttt{name} $\ulcorner{p}\urcorner$
and \texttt{coname} $\llcorner{p}\lrcorner$ of $p$, in
Fig.~\ref{fig:name-coname}.  Like with dual, we need to insert a combinator $p$
\emph{under} a quantifier.

@@

      \draw (2,0) -- (3,0) -- (3,0.5) -- (2,0.5) -- cycle;
      \node at (2.5,0.25) {\footnotesize{$p$}};

@@

      \draw (0,0) -- (1,0) -- (1,0.5) -- (0,0.5) -- cycle;
      \node at (0.50,0.25) {\footnotesize{$p^{-1}$}};

@@

>>>>

Each combinator $\rho_\bullet$ has an inverse $\rho_\bullet^{-1}$.
A rule $\mathbb{T_1}_\bullet \leftrightblackspoon

@@

Compact closed categories support various interesting constructions, such as
duals, name and coname, and trace, which eventually lead to a certain notion of
recursive higher-order functions~\cite{abramsky-2008}.  We illustrate the
expressiveness of $\Pi^/$ by defining these constructions.

%%%%%%%%%%
\subsection{\texttt{dual}}

@@

%%%
\subsection{\texttt{name} and \texttt{coname}}

Each morphism $\rho_\bullet : \order{c} \leftrightblackspoon \order{d}$ can be
given a \emph{name} and a \emph{coname}. As shown in Fig.~\ref{fig:name-coname}.
the name of $\rho_\bullet$ denoted $\ulcorner{\rho_\bullet}\urcorner$ has type
$\order{\idiso} \leftrightblackspoon \iorder{c} \circledast \order{d}$ and the
coname of $\rho_\bullet$ denoted $\llcorner{\rho_\bullet}\lrcorner$ has type
$\iorder{c} \circledast \order{d} \leftrightblackspoon \order{\idiso}$.
Intuitively, this means that for each morphism of type
$\order{c} \leftrightblackspoon \order{d}$, it is possible to construct from
``nothing'' an object $\iorder{c} \circledast \order{d}$ in the category that
denotes this morphism and dually to eliminate this object. Like with the case
for the dual construction, we need to insert a combinator \emph{under} a
quantifier.

@@

      \draw (2,0) -- (3,0) -- (3,0.5) -- (2,0.5) -- cycle;
      \node at (2.5,0.25) {\footnotesize{$\rho_\bullet$}};

@@

      \draw (0,0) -- (1,0) -- (1,0.5) -- (0,0.5) -- cycle;
      \node at (0.50,0.25) {\footnotesize{$\rho_\bullet^{-1}$}};

@@

%%%
\subsection{Higher-Order Functions}

As shown above, names and conames encode a morphism of type
$\order{c} \leftrightblackspoon \order{d}$ using morphisms between $\order{id}$
and $\iorder{c} \circledast \order{d}$. These names and conames give rise to a
notion of higher-order function that can be composed and applied. Given two
morphisms of types $\order{c} \leftrightblackspoon \order{d}$ and
$\order{d} \leftrightblackspoon \order{d}$, we can compose their names to get
the name of their sequential composition. Given the name of a morphism
$\order{c} \leftrightblackspoon \order{d}$ and given a value of type
$\order{c}$, we can produce a value of type $\order{d}$ effectively ``applying''
the morphism.
